--- 
title: "A. An Overview of AnVILBulkRNASeq"
author:
- name: Kayla Interdonato
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
- name: Martin Morgan
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output: 
    BiocStyle::html_document:
        toc: true
        toc_float: true
vignette: >
  %\VignetteIndexEntry{A_Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The goal of the `AnVILBulkRNASeq` package is to provide users all the
tools to perform a typical `DESeq2` analysis utilizing AnVIL. This
package contains all files (except for raw FASTQ files) needed to
complete the analysis.

# Installation

The following code will install necessary software for all notebooks. We start by updating installed packages to their current version, then install the 'development' version of the AnVIL package, and finally install packages used in the notebooks.

The entire process can take 20 or more minutes. Jupyter notebooks just indicate that the cell is being evaluated during this time, so it is easy to lose faith that  the command is still executing. A different approach is to start the terminal (click on `>_` in the top right), start an R session from the command line (just type `R` and then the carriage return) and copy and paste the commands below into the terminal.

```{r install, eval = FALSE}
if(!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(ask = FALSE)                   # update existing packages

pkgs = c("GenomicFeatures", "tximport", "DESeq2")   # install packages (and dependencies) for notebooks
BiocManager::install(pkgs)

BiocManager::install("Bioconductor/AnVIL")          # install the development version of the AnVIL package
```

For notebooks B and C, the AnVILBulkRNASeq needs to be installed from GitHub

```{r github, eval = FALSE}
BiocManager::install("Kayla-Morrell/AnVILBulkRNASeq")
```

# Overview

There are vignettes contained in the package that will explain each step
as well as provide code to be performed along the way.

-   *A\_Overview*: An overview of what will be accomplished in this
    workspace.
-   *B\_OrganizingData*: Downloading an example dataset and where to
    store it.
-   *C\_RunningTheWorkflow*: How to run the AnVILBulkRNASeq workflow on
    example data.
-   *D\_ManagingTheWorkflowOutput*: Obtaining the output from the
    workflow.
-   *E.0\_DESeq2Analysis*: Analyzing and visualizing the workflow output
    data.
-   *E.1\_DESeq2Analysis*: Visualizing the DESeq2 results    
-   *F\_DataCleanUp*: Once data is in the workspace, how to clean up any
    unwanted data.

Each of the above vignettes corresponds to a Jupyter notebook in the
Bioconductor-Workflow-DESeq2 workspace on AnVIL. The code will be
executable whether running from the notebooks on the workspace or
running locally from the vignettes.

# Session Information

```{r}
sessionInfo()
```
