---
title: "DESeq2 Analysis"
author:
- name: Kayla Interdonato
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
- name: Martin Morgan
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output: 
    BiocStyle::html_document:
        toc: true
        toc_float: true
vignette: >
  %\VignetteIndexEntry{E.0_DESeq2Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette will walk you through how to run a full `DESeq2` analysis
on the output data from the `AnVILBulkRNASeq` workflow. The output data
should have been retrieved in the previous vignette [Managing the
Workflow Output](D_ManagingWorkflowOutput.ipynb).

# Installation

Instructions for installing packages necessary for this notebook are given in [An Overview of AnVILBulkRNASeq](A_Overview.ipynb). Refer to
that vignette for installation steps.

Load the development version of the AnVIL package, and the packages to be used in this notebook.

```{r install}
suppressPackageStartupMessages({
    library(AnVIL)
    library(GenomicFeatures)
    library(tximport)
    library(DESeq2)
})
``` 

# Creating the `DESeq2` dataset

The files that are needed for the `DESeq2` analysis are the `quant.sf` files for
each sample. We create the path to those files (for each sample) and save them
to `files`.

```{r paths}
files_path <- paste0(getwd(), "/DRR0161%s_1/quant.sf")
files <- sprintf(files_path, 25:40)
names(files) <- sprintf("DRR0161%s", 25:40)
```

A txdb object is needed for the analysis so we download the GTF file associated
with Arabidopis thaliana and run `makeTxDbFromGFF()` on the downloaded file.

```{r txdb}
download.file("ftp://ftp.ensemblgenomes.org/pub/plants/release-28/gtf/arabidopsis_thaliana/Arabidopsis_thaliana.TAIR10.28.gtf.gz",
"Arabidopsis_thaliana.TAIR10.28.gtf.gz")
dir()

txdb <- makeTxDbFromGFF(file = "Arabidopsis_thaliana.TAIR10.28.gtf.gz")
txdb
```

We must then create a mapping between the 'TXNAME' and 'GENEID' for Arabidopsis
thaliana. This is then saved as the `tx2gene` variable for the later analysis.

```{r mapping}
k <- keys(txdb, keytype = "TXNAME")
tx2gene <- select(txdb, k, "GENEID", "TXNAME")
head(tx2gene)
```

The `tximport` function is run on the files to obtain a matrix for each sample
that contains abundance, counts, and length. This `txi` object will be used in
the later analysis.

```{r tximport}
txi <- tximport(files, type = "salmon", tx2gene = tx2gene)

tbl <- avtable('participant')
samples <- tbl[,c('participant_id', 'condition', 'type')]
samples
```

# Running `DESeq2`

Now that we have all the necessary data we are able to run the `DESeq2`
functions. `DESeqDataSetFromTximport()` will be the function used since we used
`tximport()` to obtain the necessary data.

```{r deseq2}
ddsTxi <- DESeqDataSetFromTximport(txi,
    colData = samples,
    design = ~ type + condition
)
ddsTxi
```

Now that all the data and the model design are together in a `DESeqDataSet`, we can run DESeq2's differential expression analysis:

```{r run_deseq2}
# Run DESeq2
dds <- DESeq(ddsTxi)
```

Write the output of the `DESeq()` analysis:

```{r save_deseq}
saveRDS(dds, file = "DESeq_result.RData")
AnVIL::avfiles_backup(source = "DESeq_result.RData")
```

# Sesion Information

```{r eval = TRUE}
sessionInfo()
```
